---
title: "Machine Learning Course Project"
output: html_document
---


### Summary  
  
__Using the Weight Lifting Exercise dataset, we built up a random forest machine learning model to predict type of exercise (classe) based on accelerometer data collected from belt, forearm, arm, and dumbell of 6 participants. The final model makes 100% accurate prediction in the training data and achieves 99.83% accuracy in the testing dataset.__  

### Data Preprocessing  
  
The weight lifting exercise dataset is kindly provided by the website [here](http://groupware.les.inf.puc-rio.br/har). Dataset used for this report can be downloaded [here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv).  

The following preprocessing is done:  

**1. Remove new\_window="yes" entries**  
For the new\_window variable, 19216 out of 19622 entries are marked "no" (97.93%) while only 406 marked "yes" (2.07%). The "yes" entries have values in variables like "kurtosis\_roll\_belt", "skewness\_roll\_belt" etc while the "no" entries have none. The "yes" or "no" entries do not correlate with exercise class (variable "classe"). Therefore, all entires with a "yes" value are removed from the dataset. Furthermore, the new\_window variable is removed from the dataset.  

**2. Remove missing data**  
Variables with more than half non-entry data were removed from the dataset, such as kurtosis\_roll\_belt, skewness\_roll\_belt etc.  

**3. Remove identity and time information**  
As personal identity and timestamp information is not correlated with exercise class, variables 1-5 are removed from the dataset.  

**4. Data slicing**  
The preprocessed data is saved into a variable called "trainnarrow". Using the caret and kernlab package, the dataset is sliced into a training dataset and testing dataset with the following commands:
```{r, echo=TRUE, eval=FALSE}
set.seed(1)
inTrain <- createDataPartition(trainnarrow$classe,
                               p=0.7, list=FALSE)
training <- trainnarrow[inTrain,]
testing <- trainnarrow[-inTrain,]
```

  
### Model Building and Error Evaluation  
  
As we want to make predictions to a class variable, a random forest model is selected for machine learning with train control commands to speed up the process:  
```{r, echo=TRUE, eval=FALSE}
modrf <- train(classe~., data=training, method="rf",
               trControl=trainControl(method = "cv", number = 4, allowParallel = TRUE,verboseIter = TRUE))

```

#### In sample error:  
Based on this model, we can calculate the in-sample error:
```{r, echo=TRUE, eval=FALSE}
prederr <- predict(modrf,training)
table(prederr,training$classe)
```
```{r, echo=FALSE,message=FALSE, warning=FALSE}
InEr <- matrix(c(3830,0,0,0,0,0,2603,0,0,0,0,0,2347,0,0,0,0,0,2203,0,0,0,0,0,2470),ncol=5,byrow=FALSE)
colnames(InEr) <- c("A","B","C","D","E")
rownames(InEr) <- c("A","B","C","D","E")
InEr <- as.table(InEr)
library(pander)
pandoc.table(InEr)
```
   
Therefore we make a 100% accurate prediction in the training dataset, or 0% in-sample error rate.  

#### Out of sample error:  
The out of sample error rate is predicted to be higher than the in-sample error rate, due to the over-fitting problem in building the training model. Let's calculate the out of sample error using the testing dataset:    

```{r, echo=TRUE, eval=FALSE}
pred <- predict(modrf,testing)
table(pred,testing$classe)
```
```{r, echo=FALSE,message=FALSE, warning=FALSE}
OutEr <- matrix(c(1641,0,0,0,0,2,1112,1,0,0,0,3,1002,0,0,0,0,2,942,0,0,0,0,2,1056),ncol=5,byrow=FALSE)
colnames(OutEr) <- c("A","B","C","D","E")
rownames(OutEr) <- c("A","B","C","D","E")
OutEr <- as.table(OutEr)
require(pander)
pandoc.table(OutEr)
```
  
Therefore the out of sample error rate is: (2+1+3+2+2)/5763 = 1.74%.  
The accuracy rate for prediction is: 1-1.74% = 99.83%.
  
  
### Conclusion  
  
__We built a random forest model to predict exercise class based on measured accelerometer value. The model achieves 100% accurate prediction in in-sample prediction and 99.83% accurate prediction in out-of-sample prediction. __  
  
    
      
      

